---
output:
  word_document: default
  html_document: default
---
Maternal Health Risk Data
Data has been collected from different hospitals, community clinics, maternal health cares through the IoT based risk monitoring system.
Age: Age in years when a woman is pregnant.
SystolicBP: Upper value of Blood Pressure in mmHg, another significant attribute during pregnancy.
DiastolicBP: Lower value of Blood Pressure in mmHg, another significant attribute during pregnancy.
BS: Blood glucose levels is in terms of a molar concentration, mmol/L.
HeartRate: A normal resting heart rate in beats per minute.
Risk Level: Predicted Risk Intensity Level during pregnancy considering the previous attribute
```{r}
maternal_data<-read.csv("Maternal Health Risk Data Set.csv")
str(maternal_data)
# Correcting the data type of Risk level to factor
maternal_data$RiskLevel<-as.factor(maternal_data$RiskLevel)
# Checking if the data contains missing data.
colSums(is.na(maternal_data))
# Our data set contains no missing observations.
table(maternal_data$RiskLevel)
# There are 3 risk levels females can have during pregnancy. High, mid and low risk levels. I need to design a decision tree model that will help in the classification of the risk intensity levels of pregnant females based on the available data.
```
 Using supervised learning on maternal data set to classify the  Risk Intensity Levels of pregnant mothers based on the parameters in the function.
```{r}
sample<-sample(1:nrow(maternal_data),size=0.7*nrow(maternal_data))
train_data<-maternal_data[sample,]
test_data<-maternal_data[-sample,]
library(randomForest)
library(caret)
rf_model<-randomForest(RiskLevel~.,data = train_data,proximity=T)
print(rf_model)
# The train data set model accuracy is 83.51%
test1<-predict(rf_model,test_data)
confusionMatrix(test1,test_data$RiskLevel)
# The test data set model accuracy is 85.57%
plot(rf_model)
# The plot above is used to show how the 3 risk intensity levels are distributed. 
# The plot above shows the distribution of risk intensity levels.

```
Conclusion
The classification model above is of good fit. It has an average accuracy of 84.5%. it can be used in hospitals to classify whether pregnant mothers can be classified in high, medium, and low risk intensity levels.





