---
output:
  word_document: default
  html_document: default
---
loading the data into Rstudio
```{r}
dat<-read.csv("MLData2023.csv")
library(tidyverse)
# Separate samples of non-malicious and malicious events
dat.class0 <- dat %>% filter(Class == 0) # non-malicious
dat.class1 <- dat %>% filter(Class == 1) # malicious
# Randomly select 300 samples from each class, then combine them to form a working dataset
set.seed(1234567)
rand.class0 <- dat.class0[sample(1:nrow(dat.class0), size = 300, replace = FALSE),]
rand.class1 <- dat.class1[sample(1:nrow(dat.class1), size = 300, replace = FALSE),]
# Your sub-sample of 600 observations
mydata <- rbind(rand.class0, rand.class1)
dim(mydata) # Check the dimension of your sub-sample
str(mydata)
table(mydata$Connection.State)
table(mydata$Operating.System)
table(mydata$IPV6.Traffic)
table(mydata$Ingress.Router)
```
The data types of connection states, operating systems, and IPV6.traffic  are wrong, i have to change them into factors.
Data type of variable class need to be changed into logical data type.
```{r}
mydata$Connection.State<-as.factor(mydata$Connection.State)
mydata$IPV6.Traffic<-as.factor(mydata$IPV6.Traffic)
mydata$Operating.System<-as.factor(mydata$Operating.System)
mydata$Ingress.Router<-as.factor(mydata$Ingress.Router)
mydata$Class<-factor(mydata$Class)
str(mydata)
```
Part 1
Section 1
A table for categorical and binary variables. 

```{r}
library(plyr)
p<-ldply(mydata[,c(3,6,7,9,14)],function(o) t(rbind(names(table(o)),table(o),paste0(prop.table(table(o))*100,"%"))))
colnames(p)<-c("Categorical feature","Feature","count","N %")
# Rounding to 1 decimal place
 b<-c("25.7%","9.2%","65.2%","37.3%","9.3%","0.2%","42.5%","0.2%","10.5%","70.2%","25.5%","2%","2.3%","58.7%","41.3%","50%","50%")
cbind(p,b)->p
table<-p[,-4]
colnames(table)<-c("Categorical feature","Feature","count","N %")
table
```
Section 2
summary of the Numerical variables.
```{r}
library(psych)
tab2<-describe(mydata[,c(1,2,4,5,8,10,11,12,13)])
tab2<-as.data.frame(tab2)
round(tab2,digits = 2)->tab2
tab3<-tab2[,c(5,8,9,11)]
tab3
```
Section 3
Based on section 1 above, IPV6.Traffic feature contains some invalid features.
```{r}
library(scales)
library(ggplot2)
ggplot(mydata,aes(Assembled.Payload.Size))+geom_boxplot()+ggtitle("Boxplot of Assembled.Payload.Size")
ggplot(mydata,aes(DYNRiskA.Score))+geom_boxplot()+ggtitle("Boxplot of DYNRiskA.Score")
a<-boxplot.stats(mydata$DYNRiskA.Score)$out
a
# DYNRiskA.Score contains 7 outliers.
percent(7/600)
ggplot(mydata,aes(Response.Size))+geom_boxplot()+ggtitle("Boxplot of Response.Size")
a1<-boxplot.stats(mydata$Response.Size)$out
a1
# Response.Size contains 1 outliers.
percent(1/600)
ggplot(mydata,aes(Source.Ping.Time))+geom_boxplot()+ggtitle("Boxplot of Source.Ping.Time")
a2<-boxplot.stats(mydata$Source.Ping.Time)$out
a2
#Source.Ping.Time contains 4 outliers.
percent(4/600)
ggplot(mydata,aes(Connection.Rate))+geom_boxplot()+ggtitle("Boxplot of Connection.Rate")
a3<-boxplot.stats(mydata$Connection.Rate)$out
percent(12/600)
a3
#Connection.Rate contains 12 outliers.
ggplot(mydata,aes(Server.Response.Packet.Time))+geom_boxplot()+ggtitle("Boxplot of Server.Response.Packet.Time")
ggplot(mydata,aes(Packet.Size))+geom_boxplot()+ggtitle("Boxplot of Packet.Size")
a4<-boxplot.stats(mydata$Packet.Size)$out
a4
# Packet.Size contains 4 outliers.
percent(4/600)
ggplot(mydata,aes(Packet.TTL))+geom_boxplot()+ggtitle("Boxplot of Packet.TTL")

a5<-boxplot.stats(mydata$Packet.TTL)$out
a5
# Packet.TTL contains 4 outliers.
percent(4/600)
ggplot(mydata,aes(Source.IP.Concurrent.Connection))+geom_boxplot()+ggtitle("Boxplot of Source.IP.Concurrent.Connection")
# From the above plots, variables DYNRiskA.Score,Response.Size,Source.Ping.Time,Connection.Rate, Packet.Size, and  Packet.TTL contains outliers.
```

Part 2
Section 1
Replacing outliers with NAs
```{r}
mydata[mydata$DYNRiskA.Score %in% a,"DYNRiskA.Score" ]=NA
mydata[mydata$Response.Size %in% a1,"Response.Size" ]=NA
mydata[mydata$Source.Ping.Time %in% a2,"Source.Ping.Time" ]=NA
mydata[mydata$Connection.Rate%in% a3,"Connection.Rate" ]=NA
mydata[mydata$Packet.Size %in% a4,"Packet.Size" ]=NA
mydata[mydata$Packet.TTL %in% a5,"Packet.TTL" ]=NA
colSums(is.na(mydata))
#Write to a csv file.
write.csv(mydata,"mydata.csv") 
```
Section 2
```{r}
library(scatterplot3d) #3-D plot
library(scales) #For more graphic scales options
library(tidyverse)
library(ggpubr)
library(factoextra)
mydata_numeric<-mydata[,c(1,2,4,5,8,10,11,12,13,14)]
#filter the incomplete cases
mydata_numeric_1<-na.omit(mydata_numeric)
# Scatterplot matrix to show relationship of the data
pairs(mydata_numeric_1[,-10],col=as.numeric(mydata_numeric_1$Class)+1)
```

PCA in R
Scaling is an important step before performing principal component analysis because it prevents large variable values from swaying the data even with little variability.
```{r}
pca_mydata<-prcomp(mydata_numeric_1[,-10],scale. = T)
# Summary of the data set
summary(pca_mydata)
#  Outline the individual and cumulative proportions of variance (3 decimal places) explained by each of the first 4 components.
x1<-pca_mydata$sdev
round(x1[1:4],digits = 3)
#  Outline the coefficients (or loadings) to 3 decimal places for PC1, PC2 and PC3, and describe which features (based on the loadings) are the key drivers for each of these three PCs.
loadings<-pca_mydata$rotation
as.data.frame(loadings)->loadings
loadings$PC1<-round(loadings$PC1,digits = 3)
loadings$PC2<-round(loadings$PC2,digits = 3)
loadings$PC3<-round(loadings$PC3,digits = 3)
loadings[,1:3]
library(stargazer)
```
Biplot for Pc1 and Pc2
```{r}
data1 <- data.frame(pca_mydata$x,Species=mydata_numeric_1$Class)
ggplot(data1,aes(x=PC1,y=PC2))+geom_point(aes(colour=mydata_numeric_1$Class),alpha=0.8,size=4)+theme_minimal(base_size=14)+theme(legend.position = "top")+xlab("PC1")+ylab("PC2")+ ggtitle("Biplot of PC1 and PC2")
plot(pca_mydata, type="l", main="Scree plot - Class")
# The scree plot above indicates that 2 factors should suffice. It means that the dimension have reduced from 9 to 2. The first 2 components are considered the most significant. 2 components are required to explain at least 50% of variability in the data.
fviz_pca_biplot(pca_mydata,axes = c(1,2),col.ind = mydata_numeric_1$Class,fill.ind = mydata_numeric_1$Class,legend.title=list(colour="Class",fill="Class",alpha="Class"),addEllipses = T)
#Extract the eigenvalues and proportions explained
varexp.class <- as.data.frame(summary(pca_mydata)$importance)
varexp.class
# PC1, PC2, PC3, and PC4 have the largest eigen values and the highest loading thus they can help in distinguishing malicious events. 
```
Based on the results from parts (iii) to (iv), describe which dimension (have to choose 
one) can assist with the identification of Malicious events (Hint: project all the points in 
the PCA plot to PC1 axis and see whether there is good separation between the points 
for Malicious and Non-Malicious events. Then project to PC2 axis and see if there is 
separation between Malicious and Non-Malicious events, and whether it is better than 
the projection to PC1.
```{r}
# The best dimensions to choose is 2 since it lesser that the variables present in the data set used in the analysis.
```





