---
title: "TIME SERIESTI"
output:
  word_document: default
  html_document: default
date: "2023-01-06"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(fpp3)
library(feasts)
library(forecast)
```

### Part 1
Loading the required package into Rstudio
```{r}

# First 6 rows of our data set
head(global_economy)

# Selecting Vietnam, Uruguay, and Afghanistan countries.
# Afghanistan 
Afghanistan_data <-global_economy %>% filter(Country=="Afghanistan")
ggplot(data = Afghanistan_data,aes(x=Year,y=GDP))+geom_line(col="purple")+ggtitle("GDP per capita over time for Afghanistan")
#Uruguay
Uruguay_data<- global_economy %>% filter(Country=="Uruguay")
ggplot(data =Uruguay_data,aes(x=Year,y=GDP))+geom_line(col="green")+ggtitle("GDP per capita over time for Uruguay")
#Uruguay
# Vietnam
Vietnam_data<-global_economy %>% filter(Country=="Vietnam")
ggplot(data = na.omit(Vietnam_data),aes(x=Year,y=GDP))+geom_line(col="red")+ggtitle("GDP per capita over time for Vietnam")
```

Which has highest GDP per capita?
```{r}
options(scipen = 999)
sum(Afghanistan_data$GDP,na.rm=T)
sum(Uruguay_data$GDP,na.rm=T)
sum(Vietnam_data$GDP,na.rm=T)
# Based on the above results, Vietnam has the highest GDP.
```
Checking on the above plots, it is visible that GDP has increased over the years. 


### PART 2.
Use the canadian_gas data (monthly Canadian gas production in billions of cubic metres, January 1960 – February 2005).
```{r}
str(canadian_gas)
# Canadian_gas data contains 542 observations of 2 variables.
# Plot Volume using autoplot, gg_subseries, gg_season to look at the effect of changing seasonality over time
autoplot(canadian_gas)+labs(title = "Autoplot of data")
gg_subseries(canadian_gas)+ labs(title = "subseries of data")
gg_season(canadian_gas) + labs(title = "gg_season of data")
```
Do an STL decomposition of the data. You will need to choose a seasonal window to allow for the changing shape of the seasonal component.
```{r}
decomposition<-stl(canadian_gas,s.window = "periodic")
```
How does seasonal shape change over time? [Hint: Try plotting the seasonal component using gg_season().]
```{r}
plot(decomposition,col="green")
# Under the seasonal plot above, we can say that the data exhibits a seasonal pattern

```

Can you produce a plausible seasonally adjusted series?
```{r}
sas<-as.ts(canadian_gas,frequency = 4,start=1990,end=2025)
dec_adjust<-decompose(sas,type="additive")
adjusted_volume<-sas-dec_adjust$seasonal
plot(adjusted_volume,col="green",main=" seasonally-adjusted series plot")
#plausible seasonally adjusted series analyses trends by removing all the noises present.
```



### PART 3 
Aus Retail Time Series
We will use aus_rail dataset
Using the code below, get a series (it gets a series randomly by using sample() function):
```{r}
set.seed(1234567)
myseries <- aus_retail %>%
  filter(`Series ID` == sample(aus_retail$`Series ID`,1)) 
head(myseries)
# remover NA's in the series with below:
myseries =  myseries %>% filter(!is.na(`Series ID`)) 
nrow(myseries)
# rename the column name  `Series ID` with MyRandomSeries
rename(myseries, MyRandomSeries = `Series ID`)
```
a) Run a linear regression of Turnover on its trend. Hint: use TSLM() and trend() functions)
```{r}
fit = myseries %>% model(TSLM(Turnover~ trend()))
report(fit)
# checking o the value of Multiple R-squared, it means that our model is 93.59 accurate
```
b) Forecast for next 3 years. What are the values for the next 3 years? Monthly values?
```{r}
x<-myseries$Turnover %>% forecast(h=36)
x
# In the table above Print Forecast column shows the forecasted values of the next 3 years
```
c) Plot the forecast values with original data
```{r}
ggplot(data = myseries,aes(Turnover,x$fitted))+geom_point(col="red")+geom_smooth(col="green")+ggtitle("forecast values Vs original data
")
```


d) Get the residuals, does it satisfy requirements for white noise error terms. Hint: augment() and gg_tsresiduals() functions)
```{r}
checkresiduals(x,plot = T)
# By checking the p-value and the plot above, the residuals satisfy the requirements for white noise error terms.
```


